---
title: "Homework Assignment 3-1"
author: "In this assignment we will explore methods for examining the relationship between categorical and continuous variables.<br /><br />"
output:
  bookdown::html_book
---

```{r setup,echo=FALSE,message=FALSE}
library("here")
source(here("homework","autograding","Homework_3-1.tests.R"))
.AutograderInit()
```

```{css echo=FALSE}
pre {
  margin-left: 40px;
}
div.sourceCode {
  margin-left: 40px;
}
div.solution {
  margin-left: 40px;
  border: 1px dashed red;
  padding: 2px;
}
span.solution { display: block; }
li { margin-top: 20px; }
li li { margin-top: 2px; }
```


**Walla Walla University Class Schedule**

A CSV file containing the winter 2021 class schedule at WWU has been created for your use.Load this data set into R and perform the following.

* **Problem 1**: Create a data frame named `wwu` by loading this [CSV File](https://webwork.wallawalla.edu/courses/math215/data/wwu_w21.csv) containing the winter 2021 class schedule.

```{r}
wwu <- data.frame(read.csv("https://webwork.wallawalla.edu/courses/math215/data/wwu_w21.csv"))
```


* **Problem 2**: Clean this data set by converting the credits from factors to numeric values and removing any 0-enrollment courses or courses with NA's in the credit count after conversion. Do this with the same data frame name, wwu.

```{r}
wwu <- wwu[!is.na(as.numeric(wwu$Cr)),]
wwu$Cr <- as.numeric(wwu$Cr)
wwu <- wwu[wwu$Enroll.Count > 0,]
```

```{r}
.AutogradeProblem02()
```


* **Problem 3**: Store the mean enrollment in both general studies and non-general studies courses, as indicated by the GS column, in the variables mean.GS and mean.NoGS respectively.

```{r}
mean.GS1 <- wwu[wwu$GS=="GS",]
mean.GS<- mean(mean.GS1$Enroll.Count)
mean.noGS1 <- wwu[wwu$GS!="GS",]
mean.noGS <- mean(mean.noGS1$Enroll.Count)
```

```{r}
.AutogradeProblem03()
```


* **Problem 4**: Find the standard deviation for these two groups and store them in sd.GS and sd.noGS.  

```{r}
sd.GS<- sd(mean.GS1$Enroll.Count)
sd.noGS<- sd(mean.noGS1$Enroll.Count)
```

```{r}
.AutogradeProblem04()
```


* **Problem 5**: Construct two side-by-side boxplots for this data.

```{r}
library("ggplot2")
qplot(GS,Enroll.Count, data=wwu,geom="boxplot")

```


* **Problem 6**: You should have observed two outliers in the non-general studies enrollment counts and three in the general studies enrollment counts.  Create a new data frame called wwu.noOutliers that removes these outliers.  Construct the box plots again with this new data frame.

```{r}
wwu.noOutliers <- wwu[wwu$Enroll.Count<60,]
qplot(GS,Enroll.Count, data=wwu.noOutliers,geom="boxplot")
```

```{r}
.AutogradeProblem06()
```


* **Problem 7**: Based on the comparisons of course enrollments conducted above, summarize the relationship between the variables "E.Count" and "GS".  That is, how is the enrollment count in a course at WWU this quarter related to whether that course is a general studies course or not?.

```{solution,eval=FALSE}
The general studies courses have a higher mean. The sd is also higher. This could be caused by general studies courses being open to more students.
```


* **Problem 8**: Conduct the same analysis (repeating all of the steps above) for the relationship between and Enrollment and campus (CP = College Place, BI = Billings, MI = Missoula, PD = Portland).  Use variables mean.XX and sd.XX for the mean and standard deviation, where XX is the campus code, and use wwu.noOutliers.ByCampus for the name of the data frame that has campus-specific outliers removed.

```{r}
CPEnroll <- wwu[wwu$Campus=="CP",]
mean.CP <- mean(CPEnroll$Enroll.Count)
sd.CP <- sd(CPEnroll$Enroll.Count)

BIEnroll <- wwu[wwu$Campus=="BI",]
mean.BI <- mean(BIEnroll$Enroll.Count)
sd.BI <- sd(BIEnroll$Enroll.Count)

MIEnroll <- wwu[wwu$Campus=="MI",]
mean.MI <- mean(MIEnroll$Enroll.Count)
sd.MI <- sd(MIEnroll$Enroll.Count)

PDEnroll <- wwu[wwu$Campus=="PD",]
mean.PD <- mean(PDEnroll$Enroll.Count)
sd.PD <- sd(PDEnroll$Enroll.Count)


wwu.noOutliers.byCampus <- wwu[(wwu$Campus=="CP" & wwu$Enroll.Count < 60) |
(wwu$Campus=="BI") |
(wwu$Campus=="MI") |
(wwu$Campus=="PD" & wwu$Enroll.Count < 40),]

qplot(Campus, Enroll.Count,data=wwu.noOutliers.byCampus, geom="boxplot")
```

```{r}
.AutogradeProblem08()
```


* **Problem 9**: Describe the relationship between the size of a WWU class in Winter of 2021 and the campus on which the class occurs.

```{solution,eval=FALSE}
The portland campus has the highest average. College place in general has the biggest size of classes. However the number of classes between each campus is not the same.
```


**End of Assignment**

That's the end of homework assignment 3.1.  You can now compute your total score on the autograded questions by running the code below.

```{r}
.AutograderMyTotalScore()
```