---
title: "Homework Assignment 14"
author: "In this homework assignment, we will look at examples of cleaning data sets using various techniques.<br /><br />"
output:
  bookdown::html_book
---

```{r setup,echo=FALSE,message=FALSE}
library("here")
library(forecast)
source(here("homework","autograding","Homework_14.tests.R"))
.AutograderInit()
```

```{css echo=FALSE}
pre {
  margin-left: 40px;
}
div.sourceCode {
  margin-left: 40px;
}
div.solution {
  margin-left: 40px;
  border: 1px dashed red;
  padding: 2px;
}
span.solution { display: block; }
li { margin-top: 20px; }
li li { margin-top: 2px; }
```


**Penny Dates: Asserts**

In MATH 106: Intro to Statistics, one of the labs that students must complete involves taking multiple samples of 40 pennies from a bag of several thousand pennies.  The students record the dates on each of the pennies and enter them into a spreadsheet.  Below is an actual data set recorded in one of these labs.

```{r}
df <- read.csv("https://webwork.wallawalla.edu/courses/math215/data/pennies.csv",stringsAsFactors=F)
```


1. Use the `assert` function from the `assertr` package to make the assertion that the data in the X1-X40 columns is numeric data (as we would expect years to be).  Fix any errors you find in a reasonable way, and re-verify the assertion.

```{r}
library(assertr)
#assert(df,is.numeric,colnames(df[,-1]))
df[,-1]<-as.numeric(unlist(df[,-1]))
df[232,"X35"]<-1997
```

```{r}
assert(df,is.numeric,colnames(df[,-1]))
```

```{r}
.AutogradeProblem01()
```


* **Problem 2**: Next, assert that the dates are in a reasonable range, say from 1940 through 2019.  Again, fix any errors that you find, re-verify the assertion, and explain what you did to fix the errors.

```{r}
#df  <-assert(df,within_bounds(1940,2019),colnames(df[,-1]))
#df.2<-df[,-1]
#df.remove<-df.2<1940
#df<-df[-df.remove]
```

```{r}
for(i in 2:41){
  df$X3[df$X3<1916]<-NA
}
```

```{solution,eval=FALSE}
I found the rows aren't within the range and just removed them
```


* **Problem 3**: Finally, create a chain of your two assertions using the `magrittr` package and store it in a variable `CHECKS`.  Run this on your new data frame to verify that it still passes both assertions.

```{r}
library(magrittr)
CHECKS<- . %>%
  assert(is.numeric,colnames(df[,-1])) %>%
  assert(within_bounds(1950,2019),colnames(df[,-1]))
df<-df%>%CHECKS
```

```{r}
.AutogradeProblem03()
```


**WWU Schedule: Regular Expressions**

In this section we will apply regular expressions to manipulate and normalize data values in the WWU class schedule.

```{r}
wwu <- read.csv("https://webwork.wallawalla.edu/courses/math215/data/wwu_w21.csv",stringsAsFactors=FALSE)
```


* **Problem 4**: There are several different classes that include the word Internship in their *Course.Title*.  Find their names and the total enrollment in those classes.  You will want to use the `str_subset` `str_detect` functions from the `stringr` package to do this. Save the names of the classes to the variable `internship.names` and the total enrollment to the variable `internship.enrollment` and print both out.

```{r}
library(stringr)
internship.names<-wwu$Course.Title%>% str_subset("Internship")
internship.names

internship.enrollment<-sum(wwu$Enroll.Count[wwu$Course.Title %>% str_detect("Internship")],na.rm=TRUE)
```

```{r}
.AutogradeProblem04()
```


* **Problem 5**: The `Fee` column contains any course fees.  Normalize that column so that it contains numeric values only.  If a fee is given as a per-unit value, keep that same per-unit number.  If no fee is indicated, make the fee value zero.  You may find the `str_replace` function helpful here.

```{r}
```

```{r}
.AutogradeProblem05()
```


* **Problem 6**: Several courses have a range of credits (*Cr*) available.  Replace those with the average number of credits for which the course can be taken and convert the *Cr* column to a numeric variable.

```{r}
```

```{r}
.AutogradeProblem06()
```


* **Problem 7**: Finally, calculate the number of student credit hours (number of students times number of credits) for courses that meet after 1:00 on Friday afternoons.  Save this number to `friday.afternoon.credit.hours` and print it out.

```{r}
```

```{r}
.AutogradeProblem07()
```


* **Problem 8**: Use the `group_by` and `summarize` commands from the `dplry` package to create a data frame called `wwu.total.enrolled` that has a column called `course` containing the course prefix and course number separated by a space (e.g. "MATH 215") and a second column named `enrolled` containing the total number of students enrolled in all sections of that class.

```{r}
```

```{r}
.AutogradeProblem08()
```


**End of Assignment**

That's the end of homework assignment 14.  You can now compute your total score on the autograded questions by running the code below.

```{r}
.AutograderMyTotalScore()
```
