---
title: "Homework Assignment 12"
author: "In this assignment, we will examine various ways to load data into R.<br /><br />"
output:
  bookdown::html_book
---

```{r setup,echo=FALSE,message=FALSE}
library("here")
library(forecast)
source(here("homework","autograding","Homework_12.tests.R"))
.AutograderInit()
```

```{css echo=FALSE}
pre {
  margin-left: 40px;
}
div.sourceCode {
  margin-left: 40px;
}
div.solution {
  margin-left: 40px;
  border: 1px dashed red;
  padding: 2px;
}
span.solution { display: block; }
li { margin-top: 20px; }
li li { margin-top: 2px; }
```


**Employee Data -- SQL**

The first method for loading data will be via an SQL database.  A test database has been set up for you to use on a local server. 


* **Problem 1:** Connect to the server using the `dbConnect` function from the `RMySQL` package. The IP address of the server is 172.27,8.150, the user on the server is "adminer" with password "ABC123xyz" and the database name is "employees". Save your connection object as `our.db`.

```{r}
library(RMySQL)
our.db <- dbConnect(MySQL(), user='adminer', password='ABC123xyz', dbname='employees', host='172.27.8.150')
```


* **Problem 2**: Next, let's look at the tables that are in this database. Use the `dbListTables` command to do this, save the list of tables as `our.tables` and print them out.

```{r}
our.tables<-dbListTables(our.db)
```

```{r}
.AutogradeProblem02()
```


* **Problem 3**: Next, let's list the fields (columns) that are in each table.  The first one is done for you.

```{r}
dbListFields(our.db,"current_dept_emp")
```


* **Problem 4**: Download the gender data for the rows in the *employees* table.  Use the `table` function to create a frequency distribution of gender.  Save it as `employee.gender` and print it out.

```{r}
results<-dbSendQuery(our.db,"SELECT gender FROM employees")
gender.df<-fetch(results)
dbClearResult(results)
employee.gender<-table(gender.df)
```

```{r}
.AutogradeProblem04()
```


* **Problem 5**": Next get the average salary from the *salaries* table in this database.  Save it as `mean.salary` and print it out.

```{r}
results<-dbSendQuery(our.db,"SELECT * FROM salaries")
mean.df<-fetch(results)
dbClearResult(results)
mean.salary<-mean(mean.df$salary)
```

```{r}
.AutogradeProblem05()
```


* **Problem 6**: Now construct a scatter plot of the average *salary* (y-axis) against year component of the *from_date* field (x-axis).  You can use the command `date_format(from_date,'%Y')` in your SQL query to get just the year.

```{r}
results<-dbSendQuery(our.db,"SELECT date_format(from_date,'%Y') FROM salaries")
dates.df<-fetch(results)
dbClearResult(results)
dates.vec<-unlist(dates.df)
names(mean.df)[2]<-"salary"
names(mean.df)[3]<-"from_date"
mean.df$from_date<-dates.vec
colnames(mean.df)
plot(salary~from_date,data=mean.df)

```


**iTunes -- JSON**

Another popular data format is JSON (JavaScript Object Notation).  This format is used for exchanging data seamlessly on the web.  The command below will download a list of all entities referencing the Beetles in JSON format.

```{r}
library(RCurl)
my.json <- getURL("https://itunes.apple.com/search?term=beetles&limit=1000")

```


* **Problem 7**:.  What do the contents of the `my.json` variable look like?

```{solution,eval=FALSE}
It looks like a bunch of songs and information.
```


* **Problem 8**: Use the `fromJSON` function from the `jsonlite` library to parse this into a data frame called `beatles`.  Hint: You really want the *results* element of the list that `fromJSON` creates.

```{r}
library(jsonlite)
my.list<-fromJSON(my.json)
beatles<-data.frame(my.list$results)
```

```{r}
.AutogradeProblem08()
```


* **Problem 9**: Compute the mean and variance of the track length (stored in `trackTimeMillis`) of a beatles song in seconds.  Note that times are given in milliseconds.  Save them as `beatles.mean` and `beatles.var` and print them out.

```{r}
#We talked to Professor Duncan and the API has changed causing the numbers to be off by a little bit.
beatles.mean<-mean(beatles$trackTimeMillis/1000)
beatles.var<-var(beatles$trackTimeMillis/1000)
```

```{r}
.AutogradeProblem09()
```


* **Problem 9**: Construct a histogram of the track length in seconds (note that it is given in milliseconds).

```{r}
hist(beatles$trackTimeMillis/1000)
```


* **Problem 10**: Construct side-by-side box plots of the track length in seconds for the songs in each collection, as defined by the *collectionId*.

```{r}
boxplot(trackTimeMillis/1000~collectionId,data=beatles)
```


* **Problem 11**: Apply one other data analysis technique to the data in the `beatles` data frame.  Describe what you learned from this technique.

```{r}
plot(beatles$trackTimeMillis/1000)
```

```{solution,eval=FALSE}
Answers will vary.
The scatter plot was good to visualize the times.
```


**End of Assignment**

That's the end of homework assignment 12.  You can now compute your total score on the autograded questions by running the code below.

```{r}
.AutograderMyTotalScore()
```